---
title: "activity_assignment"
author: "Supriya Sharma"
date: "3/26/2017"
output:
  html_document: default
  pdf_document: default
---
Activity assignment
==================================
## 1. Read the activity report
```{r}
data <- read.csv(file="activity.csv", header = T, sep=",")
```

## 2. Histogram of the total number of steps taken each day
```{r}
steps <-aggregate(data$steps, by= list(data$date),FUN = sum, na.rm=FALSE)
hist(steps$x, 
     main= "Total number of steps taken each day", 
     xlab= "steps each day", 
     col= "green")
```

## 3. Mean and median number of steps taken each day
```{r}
mean_steps <- mean(steps$x, na.rm=T)
mean_steps
median_steps <- median(steps$x, na.rm=T)
median_steps
```

## 4. Time series plot of the average number of steps taken
```{r}
steps_int <-aggregate(data$steps, by= list(data$interval),FUN = mean, na.rm=TRUE)
plot(x= steps_int[,1],
     y= steps_int[,2],
     main= "Average number of steps taken per interval", 
     type = "l",
     col="blue", 
     xlab= "Time interval", 
     ylab ="steps")
```


## 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
int_maximum_steps <- steps_int[which.max(steps_int[,2]),1]
int_maximum_steps
```
       
## 6. Code to describe and show a strategy for imputing missing data
```{r}
missing_NA <- sum(is.na(data$steps))

## combined activity data with steps interval to eventually calculate mean
data_imputated <- cbind(data, steps_int[,2])
names(data_imputated)[4] <- c("mean")
data_imputated$steps <- ifelse( is.na(data_imputated$steps), data_imputated$mean, data_imputated$steps)
```

## 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
## total number of steps per day in imputated data
data_imputated_stepsPerDay <- aggregate(data_imputated$steps, by=list(data_imputated$date), FUN=sum)

## mean of total number of steps per day in imputated data
data_imputated_stepsPerDayMean <- mean(data_imputated_stepsPerDay$x)
data_imputated_stepsPerDayMean

## median of total number of steps per day in imputated data
data_imputated_stepsPerDayMedian <- median(data_imputated_stepsPerDay$x,)
data_imputated_stepsPerDayMedian

hist(data_imputated_stepsPerDay$x, 
     breaks=20,
     main="Total number of steps take per day", 
     col="gold", 
     xlab="Steps taken")
```

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
data_imputated$date <- strptime(data_imputated$date, "%Y-%m-%d")
data_imputated$weekend <- (weekdays(data_imputated$date) %in% c("Sunday", "Saturday"))

data_weekend <-data_imputated[data_imputated$weekend == TRUE,]
data_weekday <- data_imputated[data_imputated$weekend == FALSE,]

## average of steps taken on weekends and weekdays
av_weekend_steps <- aggregate(data_weekend$steps, by=list(data_weekend$interval), FUN=mean)
av_weekday_steps <- aggregate(data_weekday$steps, by=list(data_weekday$interval), FUN=mean)

par(mfrow=c(2,1))

plot( x=av_weekend_steps[,1], 
      y=av_weekend_steps[,2], 
      type= "l",
      col="purple",
      main="Weekend average steps per interval",
      ylab="Average steps", 
      xlab="Interval ID")

plot( x=av_weekday_steps[,1], 
      y=av_weekday_steps[,2],
      type= "l",
      col="magenta",
      main="Weekday average steps per interval",
      ylab="Average steps", 
      xlab="Interval ID")
```

